---
title: Basic Statistical Calculations
---

In this chapter we will learn how to carry out some basic statistical measurements in R.
The measurements are-

+ Mean
+ Variance
+ Median
+ Range
+ Co-variance
+ Correlation

The formula for the measurements are

$$ Mean= \frac{1}{n} \sum_{i=1}^{n} x_i$$
$$Variance = \frac{1}{n}\sum_{i=1}^{n} (x_i-\bar{x})$$
$$ Median = l+\dfrac{\frac{N}{2}-f_c}{f}* h$$

$$Range = Maximum \ \   observation - Minimum \ \ \   observation$$
$$ Covariance = \frac{1}{n}\sum_{i=1}^{n}x_iy_i- \bar{x}\bar{y}$$
$$ Correlation = \dfrac{Cov(x,y)}{\sqrt{Variance(x)}\sqrt{Variance(y)}}$$
Let us consider the following data
```{r}
weight=c(45,81,47,58,61,76,47,44,66,55)
height=c(140,165,145,160,166,170,156,161,159,163)
```

Once the data are being entered in R, we can now perform various statistical calculations:

### Mean

Generally by **`mean`** we calculate **`Arithmetic mean`**. Arithmetic mean gives the center of our data

```{r}
mean_weight=mean(weight)
mean_weight
mean_height=mean(height)
mean_height
```

### Variance

Variance gives the amount of variation present in our data. Or how much our information are scattered from the center of our data

```{r}
variance_weight=var(weight)
variance_weight
variance_height=var(height)
variance_height
```

### Standard deviation

Standard deviation is simply the square root of variance

```{r}
sd_weight=sd(weight)
sd_weight
sd_height=sd(height)
sd_height
```


### Median

Median is the mid point of our information when arranged in ascending or descending order.

```{r}
median_weight=median(weight)
median_weight

median_height=median(height)
median_height
```

### Range

Range gives the minimum and maximum values of the data.

```{r}
range_weight=range(weight)
range_weight

range_height=range(height)
range_height

```

### Covariance

Co-variance is a measurement of variation present in two sets of data when considered together. In our example, we are calculating how weight and height varies together.

```{r}
cov_weight_height=cov(weight, height)
cov_weight_height
```

### Correlation

Correlation is the measurement of association or bonding between two variables. In this example, correlation will tell us what kind of relationship weight and height posses.

```{r}
correlation=cor(weight, height)

correlation
```

+ If the value of correlation is greater than 0, then we say that the two variables are positively correlated.

+ If the value of correlation is less than 0, then we say that the two variables are negatively correlated.

+ If the value of correlation is 0, then we say that the two variables are not correlated.

In our example the correlation is `r correlation` which is greater than 0, and hence we say that weight and heights are positively correlated.