---
title: Vectors
---

A vector is a variable which can store more than one value under the same variable.

```{r}
a=c(1,2,3,4,5,6,7,8,9,10)
a      # This will display the values stored in variable "a"
sum(a) # This will give the total sum of all the elements of "a"
a^2  # This will square each and every elements of the vector "a"
b=c(2,3,4,5,6,7,8,9,10,11)
a+b  # This will add corresponding elements of the two vectors "a" and "b" 
b-a  # This will give the diffrence between corresponding elements of "b" and "a"
a*b  # This will multiply corresponding elements of "a" and "b".

a/b  # This will divide the elements of vector "a" by 
     # corresponding elements of vector "b".

```



## Arranging a vector in increasing or decreasing order

```{r}
aa=c(7,1,9,4,7,3,5,8,41,36,74,2)
sort(aa)
sort(aa, decreasing = T)
```


## Some other in-built functions for vectors

The `seq` function is used to create a sequence of numbers

```{r}
a_seq = seq(from=0, to= 10, by=1)
a_seq
```

  + from = first term
  + to = last term
  + by = the difference between two numbers

The `rep` function can be use to generate a vector with repeated value.

```{r}
a_rep=rep(5, times=10)
a_rep
```
Here a vector is created with the value 5 repeated 10 times.